Chang, Agrawal, and Cameron survey the power management techniques employed by Android and how popular applications abuse or take advantage of those techniques\cite{energy-aware}.
They explore the idea software written for mobile devices to be optimized for efficient power optimization and in many cases, the application developer doesn't necessarily take power into
consideration.
They do several case studies of popular applications and analyze how they interact with the CPU, the screen, and other aspects of the mobile devices. 
They came to the conclusion that even though Android does a great job to aggresively save power, application developers needs to create their software with power consumption in mind.
The paper is a beneficial overview of the Android power management system as well as several tools that can be used to monitor the power usage of a device's underlying hardware.

Pathak, Hu, and Zhang help application developers create power conserving software through their energy profiling tool called \emph{eprof}\cite{eprof}.
The tool allows developers to monitor power consumption with multiple granularities and determine which pieces are code consume the most energy.
This is useful for catching programming bugs, such as holding onto a wakelock too long, as well as directing attention to code that needs to be optimized.
It also illuminates the fact that large amounts of energy are used in I/O operations. 
This implies that in order to save energy it is important to only use 3G, GPS, and WiFi when it is truly necessary.
Therefore, it is clear that disabling these aspects of a mobile device can make profound differences in battery life.

\begin{thebibliography}{9}

\bibitem{eprof}
Abhinav Pathak, Y. Charlie Hu, and Ming Zhang,
  ``Fine Grained Energy Accounting on Smartphones with Eprof,"
  in \emph{Proc. of EuroSys}, 2012. \\ 
  URL: \verb https://engineering.purdue.edu/~ychu/publications/eurosys12_eprof.pdf .

\bibitem{energy-aware}
Hung-Ching Chang, Abhishek R Agrawal, and Kirk W Cameron,
  ``Energy-Aware Computing for Android Platforms," \\
  URL: \verb http://scape.cs.vt.edu/wp-content/uploads/2012/08/ITJ12_Android_Energy-Aware.pdf .

\end{thebibliography}


